---
title: "SP group project renewal  process"
output: word_document
---

```{r}
#(a)
#for uniform distribution 
set.seed(345)
draws <- runif(20)
draws
```

Drawing a random sample of n=20 from U(0,1)

```{r}
find_n <- function(u) {
  which(cumsum(u) > 1)[1]  # which partial sum first exceeds 1?
}

n <- find_n(draws)
draws[1:n]
```
We have defined a function to check which partial sum exceeds 1 with the help of cumsum
After that we have shown result for which set of RV's the cumulative sum is exceeding 1

```{r}
sample_from_N <- function(m)
{
  draws <- matrix(runif(m * 20), ncol = 20)
  n.many <- apply(draws, 1, find_n)
  estimated.probs <- table(factor(n.many, levels = 2:10)) / m  
  return(estimated.probs)
}
m <- 1000
approx.N <- sample_from_N(1000)  # approximate prob. dist'n of N
approx.N

```

From the table we can see that the probability associated with each RV's sum is given, i.e., The probability for the RV's to cross a sum of 2 is equal to 0.489 and so on for every cumulative sum


```{r}
#for log-normal N(0,1)
set.seed(65)
draws <- rlnorm(20)
draws
```

Drawing a random sample of n=20 from Log-normal(0,1)

```{r}
find_n <- function(u) {
  which(cumsum(u) >= 1)[1]  # which partial sum first exceeds 1?
}

n <- find_n(draws)
draws[1:n]
```
We have defined a function to check which partial sum exceeds 1 with the help of cumsum
After that we have shown result for which set of RV's the cumulative sum is exceeding 1

```{r}
sample_from_N <- function(m)
{
  draws <- matrix(rlnorm(m * 20), ncol = 20)
  n.many <- apply(draws, 1, find_n)
  estimated.probs <- table(factor(n.many, levels = 1:10)) / m  
  return(estimated.probs)
}
m <- 1000
approx.N <- sample_from_N(1000)  # approximate prob. dist'n of N
approx.N
```
From the table we can see that the probability associated with each RV's sum is given, i.e., The probability for the RV's to cross a sum of 1 is equal to 0.504 and so on for every cumulative sum.

```{r}
#(b)
time=readline(prompt = "Enter any time between 2 buses =")
```

Using the readline function to read a line from the terminal , i.e. prompting the user for input.

```{r}
t1=c(0,time)
print(paste("The expected time for next bus to come is",mean(as.numeric(t1)),"mins"))
```
Setting the range for time and then calculating the mean to get the expected time for arrival for the next bus (this is the wrong inter-arrival time).

```{r}
#Using Inspection paradox
#Assuming 5 buses to come in first "time" mins and we do it for an hour(60 mins)
t2=readline(prompt = "What is the time interval you want the buses to come in your given time between 2 buses\n")
c1=as.numeric(t2)
inter=c(c1,c1,c1,c1,c1,(60-as.numeric(time)))
avgarrival_time_between_buses=mean(inter)
avgarrival_time_between_buses
```
In the similar manner, we have assumed 5 buses to come in first "t" mins and then we do it for an hour . The average arrival time between buses comes out to be 8.333 mins. 

```{r}
#using the inspection paradox formula
avgtime_wait=mean(as.numeric(t1)*((c1/60)*(c1/2))+((60-as.numeric(time))/60)*(60-as.numeric(time))/2)
avgtime_wait
```
Using the inspection paradox formula , the actual waiting time for the arrival of 2 buses is equal to 13.667 which is the actual waiting time unlike 8.333

```{r}
#(c)
#for U(0,1)
ren.process <- function(a,b,n){
  s<-numeric(n+1)
  # set S_0 = 0
  s[1] <-0
  # generate vector of iid Uniform random variables:
  set.seed(45)
  x <-replicate(n,runif(1,a,b))
  # assign wait time to vector s in for loop:
  for (k in 1:n){
    
    s[k+1] <-sum(x[1:k])
    
  }
  # return vector of wait time
  return(s)
}
```

First, we initialize the vector of total wait time for the arrival of each event then we set the vector s as 0. Now we generate a vectors of iid uniform U(0,1) random variables to make this process a renewal process.

```{r}
a<-0
b<-1
#For uniform U(0,1)
s_list1 <-ren.process(a,b,20)
s_list1
s_list2 <-ren.process(a,b,50)
s_list2
s_list3 <-ren.process(a,b,100)
s_list3
s_list4 <-ren.process(a,b,1000)
s_list4
s_list5 <-ren.process(a,b,10000)
s_list5
```

```{r}
#for N(5,9)
ren.process1 <- function(a1,b1,n){
  
  # initialize vector of total wait time for the arrival of each event:
  s<-numeric(n+1)
  # set S_0 = 0
  s[1] <-0
  # generate vector of iid Normal random variables:
   set.seed(135)
  x <-replicate(n,rnorm(1,a1,b1))
  # assign wait time to vector s in for loop:
  for (k in 1:n){
    
    s[k+1] <-sum(x[1:k])
    
  }
  # return vector of wait time
  return(s)
  
}
a1<-5
b1<-9
#For normal N(5,9)
s_list6 <-ren.process1(a1,b1,20)
s_list6
s_list7 <-ren.process1(a1,b1,50)
s_list7
s_list8 <-ren.process1(a1,b1,100)
s_list8
s_list9 <-ren.process1(a1,b1,1000)
s_list9
s_list10 <-ren.process1(a1,b1,10000)
s_list10
```

```{r}
#(d)
#for U(0,1)
par(mfrow=c(3,2))
plot(stepfun(0:(20-1), s_list1), 
     do.points = TRUE,
     pch = 16,
     col.points = "red",
     verticals = FALSE,
     main = 'Realization of a Renewal process with parameters of Uniform distribution',
     xlab = 'Time of arrival',
     ylab = 'Number of arrivals')
plot(stepfun(0:(50-1), s_list2), 
     do.points = TRUE,
     pch = 16,
     col.points = "red",
     verticals = FALSE,
     main = 'Realization of a Renewal process with parameters of Uniform distribution',
     xlab = 'Time of arrival',
     ylab = 'Number of arrivals')
plot(stepfun(0:(100-1), s_list3), 
     do.points = TRUE,
     pch = 16,
     col.points = "red",
     verticals = FALSE,
     main = 'Realization of a Renewal process with parameters of Uniform distribution',
     xlab = 'Time of arrival',
     ylab = 'Number of arrivals')
plot(stepfun(0:(1000-1), s_list4), 
     do.points = TRUE,
     pch = 16,
     col.points = "red",
     verticals = FALSE,
     main = 'Realization of a Renewal process with parameters of Uniform distribution',
     xlab = 'Time of arrival',
     ylab = 'Number of arrivals')
plot(stepfun(0:(10000-1), s_list5), 
     do.points = TRUE,
     pch = 16,
     col.points = "red",
     verticals = FALSE,
     main = 'Realization of a Renewal process with parameters of Uniform distribution',
     xlab = 'Time of arrival',
     ylab = 'Number of arrivals')


```

```{r}
#for N(5,9)
par(mfrow=c(3,2))
plot(stepfun(0:(20-1), s_list6), 
     do.points = TRUE,
     pch = 16,
     col.points = "red",
     verticals = FALSE,
     main = 'Realization of a Renewal process with parameters of Normal distribution',
     xlab = 'Time of arrival',
     ylab = 'Number of arrivals')
plot(stepfun(0:(50-1), s_list7), 
     do.points = TRUE,
     pch = 16,
     col.points = "red",
     verticals = FALSE,
     main = 'Realization of a Renewal process with parameters of Normal distribution',
     xlab = 'Time of arrival',
     ylab = 'Number of arrivals')
plot(stepfun(0:(100-1), s_list8), 
     do.points = TRUE,
     pch = 16,
     col.points = "red",
     verticals = FALSE,
     main = 'Realization of a Renewal process with parameters of Normal distribution',
     xlab = 'Time of arrival',
     ylab = 'Number of arrivals')
plot(stepfun(0:(1000-1), s_list9), 
     do.points = TRUE,
     pch = 16,
     col.points = "red",
     verticals = FALSE,
     main = 'Realization of a Renewal process with parameters of Normal distribution',
     xlab = 'Time of arrival',
     ylab = 'Number of arrivals')
plot(stepfun(0:(10000-1), s_list10), 
     do.points = TRUE,
     pch = 16,
     col.points = "red",
     verticals = FALSE,
     main = 'Realization of a Renewal process with parameters of Normal distribution',
     xlab = 'Time of arrival',
     ylab = 'Number of arrivals')
```











